---
title: Experimenting with Ct values
---



```{r}
library(tidyverse)
tpr1 <- 0.983
tpr10 <- 0.8
est_rct <- function(r0, ctmax, E)
{
	r0 * (1 + E)^ctmax
}

est_r0 <- function(ct, E, rct)
{
	rct / (1+E)^ct
}

est_ct <- function(r0, E, rct)
{
	log(rct/r0) / log(1+E)
}
```

If Ct has some distribution, what distribution must E have such that 10x dilution drops sensitivity from 0.983 to 0.8?

```{r}
ct_fun <- function(n, ct_alpha, ct_beta, max_ct, min_ct)
{
	rbeta(n, ct_alpha, ct_beta) * (max_ct - min_ct) + min_ct
}

get_E <- function(uci, lci, perc=0.99)
{
	mu <- mean(c(uci, lci))
	s <- (mu - lci) / qnorm(perc)
	print(mu)
	print(s)
	rnorm(100000, mu, s)
}

n <- 100000
rct <- 10
ctthresh <- 35
e_max <- 0.9
e_min <- 0.4
e_alpha <- 2
e_beta <- 4
sensitivity_1 <- 0.983
max_ct <- 42
min_ct <- 12
o <- optim(par = c(3,4), fn = function(param, min_ct, max_ct, sensitivity, n, ctthresh)
{
	ct <- ct_fun(n, param[1], param[2], max_ct, min_ct)
	(sum(ct > ctthresh)/n - sensitivity)^2
}, min_ct=min_ct, max_ct=max_ct, sensitivity=1-sensitivity_1, n=n, ctthresh=ctthresh)
ctd <- ct_fun(n, o$par[1], o$par[2], ct_max, ct_min)
Ed <- rbeta(n, e_alpha, e_beta) * (e_max-e_min) + e_min
r0 <- est_r0(ctd, Ed, rct)
r0_10 <- r0 / 10
ctd_10 <- est_ct(r0_10, Ed, rct)
sum(ctd_10 < ctthresh) / n
par(mfrow=c(2,2))
hist(ctd)
hist(ctd_10)
hist(Ed)
hist(r0)
```

Problem - E needs to be too low. Try making E and Ct correlated

```{r}
library(mvtnorm)

get_ct <- function(n, uci, lci, perc=0.99)
{
	mu <- mean(c(uci, lci))
	s <- (mu - lci) / qnorm(perc)
	rnorm(100000, mu, s) %>% pmax(., 1)
}

get_E <- function(n, uci, lci, perc=0.99)
{
	mu <- mean(c(uci, lci))
	s <- (mu - lci) / qnorm(perc)
	rnorm(100000, mu, s) %>% pmax(., 0) %>% pmin(., 1)
}

n <- 100000
rct <- 10
ctthresh <- 35
e_max <- 0.9
e_min <- 0.4
e_alpha <- 2
e_beta <- 4
sensitivity_1 <- 0.983
max_ct <- 42
min_ct <- 12
o <- optim(par = c(42,12), fn = function(param, sensitivity, n, ctthresh)
{
	ct <- get_ct(n, param[1], param[2])
	(sum(ct > ctthresh)/n - sensitivity)^2
}, sensitivity=1-sensitivity_1, n=n, ctthresh=ctthresh)
o
ctd <- get_ct(n, o$par[1], o$par[2])
Ed <- get_E(n, 0.9, 0.65)
init <- rmvnorm(n, c(mean(Ed), mean(ctd)), sigma=matrix(c(var(Ed), -0.05, -0.05, var(ctd)), 2, 2))
Ed <- init[,1] %>% pmax(., 0) %>% pmin(., 1)
ctd <- init[,2] %>% pmax(., 1)
r0 <- est_r0(ctd, Ed, rct)
r0_10 <- r0 / 2
ctd_10 <- est_ct(r0_10, Ed, rct)
cor(init)^2
sum(ctd < ctthresh) / n
sum(ctd_10 < ctthresh) / n
par(mfrow=c(2,2))
hist(ctd)
hist(Ed)
hist(ctd_10)
hist(r0, breaks=100)

plot(ctd, Ed)

```

This doesn't solve it. Try making Ct flatter - that way small change in viral loads will push more of the distribution over Ct threshold of 35

```{r}
get_ct <- function(n, uci, lci, ct_alpha=1.1, ct_beta=1.1)
{
	rbeta(n, ct_alpha, ct_beta) * (uci-lci) + lci
}

get_E <- function(n, uci, lci, perc=0.99)
{
	mu <- mean(c(uci, lci))
	s <- (mu - lci) / qnorm(perc)
	rnorm(100000, mu, s) %>% pmax(., 0) %>% pmin(., 1)
}

n <- 100000
rct <- 10
ctthresh <- 35
e_max <- 0.9
e_min <- 0.65
e_alpha <- 2
e_beta <- 4
sensitivity_1 <- 0.983

o <- optim(par = c(42,12), fn = function(param, sensitivity, n, ctthresh)
{
	ct <- get_ct(n, param[1], param[2])
	(sum(ct > ctthresh)/n - sensitivity)^2
}, sensitivity=1-sensitivity_1, n=n, ctthresh=ctthresh)
o
ctd <- get_ct(n, o$par[1], o$par[2])
Ed <- get_E(n, e_max, e_min)
r0 <- est_r0(ctd, Ed, rct)
r0_10 <- r0 / 10
ctd_10 <- est_ct(r0_10, Ed, rct)
cor(init)^2
sum(ctd < ctthresh) / n
sum(ctd_10 < ctthresh) / n
par(mfrow=c(2,2))
hist(ctd)
hist(Ed)
hist(ctd_10)
hist(r0, breaks=100)
```

That is much better. Now optimise to identify parameters of Ct and E distributions

```{r}
set.seed(12345)
n <- 500000
rct <- 10
ctthresh <- 35
e_max <- 0.9
e_min <- 0.65
e_alpha <- 2
e_beta <- 4
sensitivity_1 <- 0.983
tpr1 <- 0.983
tpr10 <- 0.8

fn <- function(param, n, rct, e_max, e_min, ctthresh, tpr1=tpr1, tpr10=tpr10)
{
	ctd <- get_ct(n, param[1], param[2], param[3], param[4])
	Ed <- rbeta(n, param[5], param[6]) * (e_max-e_min) + e_min
	r0 <- est_r0(ctd, Ed, rct)
	r0_10 <- r0 / 10
	ctd_10 <- est_ct(r0_10, Ed, rct)
	s1 <- sum(ctd < ctthresh)/n
	s10 <- sum(ctd_10 < ctthresh)/n
	return((s1 - tpr1)^2 + (s10 - tpr10)^2)
}

(efficiency_params <- optim(par = c(35, 15, 1.1, 1.1, 1, 1), fn=fn, n=n, rct=rct, e_max=e_max, e_min=e_min, ctthresh=ctthresh, tpr1=tpr1, tpr10=tpr10, control=list(maxit=1000)))

ctd <- get_ct(n, efficiency_params$par[1], efficiency_params$par[2], efficiency_params$par[3], efficiency_params$par[4])
Ed <- rbeta(n, efficiency_params$par[5], efficiency_params$par[6]) * (e_max-e_min) + e_min
r0 <- est_r0(ctd, Ed, rct)
r0_10 <- r0 / 10
ctd_10 <- est_ct(r0_10, Ed, rct)
s1 <- sum(ctd < ctthresh)/n
s10 <- sum(ctd_10 < ctthresh)/n

sum(ctd < ctthresh) / n
sum(ctd_10 < ctthresh) / n

par(mfrow=c(2,2))
hist(ctd)
hist(Ed)
hist(ctd_10)
hist(r0, breaks=100)
```

Change in sensitivity over dilutions


```{r}
dat <- tibble(
	D=seq(1,100), 
	sensitivity=sapply(D, function(x) 
	{
		r0_x <- r0 / x
		ctd_x <- est_ct(r0_x, Ed, rct)
		sum(ctd_x < ctthresh)/n
	})
)
plot(sensitivity ~ D, dat)


```

