---
title: Lateral Flow Device sensitivity model
---

To compare the pooling model against lateral flow results, see the following document:

[Preliminary report from the Joint PHE Porton Down & University of Oxford SARS-CoV-2 test development and validation cell](https://www.ox.ac.uk/sites/files/oxford/media_wysiwyg/UK%20evaluation_PHE%20Porton%20Down%20%20University%20of%20Oxford_final.pdf)

Figure 2 provides a relationship between Cq value from PCR tests of known positive samples, and LFD sensitivity. Through visual inspection, can obtain the following approximate points:


```{r}
library(dplyr)
d <- tribble(
	~cq, ~sensitivity,
	18.0, 1,
	19.0, 0.98,
	23.0, 0.96,
	26.5, 0.89,
	29.5, 0.65,
	32.5, 0.32,
	34.5, 0.22,
)
plot(d)
```

Fit a sigmoid model to this relationship, to be used to infer the probability of testing positive in LFD based on modelled Ct value

```{r}
lfd_fit <- nls(sensitivity ~ SSlogis(cq, Asym, xmid, scal), data = d)
summary(lfd_fit)
plot(seq(10, 50, length.out = 100), 
      predict(lfd_fit, newdata = data.frame(cq = seq(10, 50, length.out = 100))),
      xlab="RT-qPCR Cq value", ylab="LFD sensitivity", type="l")
points(d, col="red", pch=19)
```

This model is:

$$
sensitivity = Asym / (1 + exp((xmid - cq)/scal))
$$

Save model fit

```{r}
lfd_fit <- summary(lfd_fit)
save(lfd_fit, file="../data/lfd_fit.rdata")
```

